---
title: "G7 model"
author: "Xiwei Zhang"
date: "2026-03-07"
output: 
  pdf_document:
    latex_engine: xelatex
---

```{r prepare}
library(VGAM)
library(dplyr)
library(tidyr)
library(readr)
library(readxl)
library(stringr)
library(lubridate)
library(zoo)
```

```{r country codes}
code_to_country <- c(
  AR = "Argentina",
  AU = "Australia",
  BR = "Brazil",
  CA = "Canada",
  CN = "China",
  DE = "Germany",
  FR = "France",
  GB = "United Kingdom",
  ID = "Indonesia",
  IN = "India",
  IT = "Italy",
  JP = "Japan",
  KR = "Korea, Republic of",
  MX = "Mexico",
  RU = "Russian Federation",
  SA = "Saudi Arabia",
  TR = "Turkey",
  US = "United States",
  ZA = "South Africa",
  XM = "Other_ECB"
)
```

```{r build 5-anchor composition  }
topic <- read_csv(
  "data/g20bis_monthly_topic_proportions.csv",
  show_col_types = FALSE
) %>%
  mutate(
    date = as.Date(date),
    date_m = floor_date(date, "month"),
    anchor5 = if_else(anchor %in% c("market", "bank", "growth", "govern"),
                      anchor, "other"),
    avg_proportion = as.numeric(avg_proportion)
  ) %>%
  group_by(country, date_m, anchor5) %>%
  summarise(
    avg_proportion = sum(avg_proportion, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = anchor5,
    values_from = avg_proportion,
    values_fill = 0
  ) %>%
  filter(date_m >= as.Date("2010-01-01"))

## quick check
print(names(topic))
print(head(topic))
```

```{r real GDP convert to monthly}
real_m <- read_excel(
  "data/macro_control_vars.xlsx",
  sheet = "realGDP"
) %>%
  rename(date = 1) %>%
  pivot_longer(-date, names_to = "series", values_to = "realGDP") %>%
  mutate(
    date = as.Date(date),
    code = str_extract(series, "[A-Z]{2,3}$"),
    country = recode(code, !!!code_to_country),
    q = as.yearqtr(date),
    q_start = as.Date(q, frac = 0)
  ) %>%
  filter(!is.na(country)) %>%
  group_by(country, q, q_start) %>%
  summarise(
    realGDP = mean(realGDP, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  rowwise() %>%
  mutate(
    date_m = list(seq.Date(q_start, by = "month", length.out = 3))
  ) %>%
  unnest(date_m) %>%
  ungroup() %>%
  select(country, date_m, realGDP)
```

```{r Read the other 3 macro variables as monthly data}
read_monthly_sheet <- function(sheet_name) {
  read_excel(
    "data/macro_control_vars.xlsx",
    sheet = sheet_name
  ) %>%
    rename(date = 1) %>%
    pivot_longer(-date, names_to = "series", values_to = "value") %>%
    mutate(
      date = as.Date(date),
      date_m = floor_date(date, "month"),
      code = str_extract(series, "[A-Z]{2,3}$"),
      country = recode(code, !!!code_to_country)
    ) %>%
    filter(!is.na(country)) %>%
    group_by(country, date_m) %>%
    summarise(value = mean(value, na.rm = TRUE), .groups = "drop") %>%
    rename(!!sheet_name := value)
}

headline_m <- read_monthly_sheet("headlineInflation")
usd_m      <- read_monthly_sheet("usdExchangeRate")
policy_m   <- read_monthly_sheet("policyRate")

# 5. Build macro panel
macro_panel <- real_m %>%
  full_join(headline_m, by = c("country", "date_m")) %>%
  full_join(usd_m,      by = c("country", "date_m")) %>%
  full_join(policy_m,   by = c("country", "date_m")) %>%
  filter(date_m >= as.Date("2010-01-01"))

print(head(macro_panel))

final_table <- inner_join(
  topic,
  macro_panel,
  by = c("country", "date_m")
) %>%
  arrange(country, date_m)

print(final_table, width = Inf, n = 20)

```

```{r Dirichlet regression }
comp_cols <- c("bank", "govern", "growth", "market", "other")

df2 <- final_table %>%
  select(
    country, date_m, all_of(comp_cols),
    realGDP, headlineInflation, usdExchangeRate, policyRate
  ) %>%
  na.omit()
```

```{r G7 model preapre }
comp_cols <- c("bank","govern","growth","market","other")

df2 <- final_table %>%
  select(country, date_m, all_of(comp_cols),
         realGDP, headlineInflation, usdExchangeRate, policyRate) %>%
  na.omit()

df2[comp_cols] <- lapply(df2[comp_cols], as.numeric)
S <- rowSums(df2[, comp_cols])
df2[, comp_cols] <- df2[, comp_cols] / S

eps <- 1e-3
K <- length(comp_cols)
df2[, comp_cols] <- (df2[, comp_cols] + eps) / (1 + K * eps)

delta <- 0.002
df2[, comp_cols] <- (1 - delta) * df2[, comp_cols] + delta * (1 / K)

g7_countries <- c(
  "Canada",
  "France",
  "Germany",
  "Italy",
  "Japan",
  "United Kingdom",
  "United States"
)

df2_10y <- df2 %>%
  dplyr::filter(
    country %in% g7_countries,
    date_m >= as.Date("2014-01-01"),
    date_m <= as.Date("2024-01-01")
  ) %>%
  droplevels() %>%
  dplyr::mutate(
    country = factor(country),
    timeIndex = (as.integer(format(date_m, "%Y")) - 2015) * 12 +
      as.integer(format(date_m, "%m")) + 1,
    timeIndex_s = as.numeric(scale(timeIndex))
  )

df2_10y$headlineInflation_s <- as.numeric(scale(df2_10y$headlineInflation))
df2_10y$policyRate_s        <- as.numeric(scale(df2_10y$policyRate))
df2_10y$usdExchangeRate_s   <- as.numeric(scale(df2_10y$usdExchangeRate))
df2_10y$realGDP_s           <- as.numeric(scale(df2_10y$realGDP))
df2_10y$country             <- droplevels(factor(df2_10y$country))

```

```{r G7 model }

# G7 single-country models: final version


# 1) Response adjustment
comp <- c("bank", "govern", "growth", "market", "other")
y <- as.matrix(df2_10y[, comp])

K <- ncol(y)

y_adj <- y
y_adj[y_adj < 0] <- 0
y_adj <- y_adj / rowSums(y_adj)

eps_resp <- 1e-3
y_adj <- (y_adj + eps_resp) / (1 + K * eps_resp)

delta_resp <- 0.002
y_adj <- (1 - delta_resp) * y_adj + delta_resp * (1 / K)

y_adj <- y_adj / rowSums(y_adj)
storage.mode(y_adj) <- "double"

## 2) Safe scaling helper
safe_scale <- function(x) {
  x <- as.numeric(x)
  s <- sd(x, na.rm = TRUE)
  m <- mean(x, na.rm = TRUE)

  if (!is.finite(s) || s == 0) {
    return(rep(0, length(x)))
  }

  (x - m) / s
}

# 3) Fit main countries except US and Japan
#    Italy uses ps.int = 4
#    others use ps.int = 5

g7_models <- list()

main_countries <- c("Canada", "France", "Germany", "Italy", "United Kingdom")

for (cty in main_countries) {
  cat("\n====================\n")
  cat("Fitting:", cty, "\n")

  idx <- which(df2_10y$country == cty)
  ord <- order(df2_10y$date_m[idx])

  dat_cty <- df2_10y[idx[ord], ]
  y_cty   <- y_adj[idx[ord], , drop = FALSE]

  dat_cty$headlineInflation_sc <- safe_scale(dat_cty$headlineInflation)
  dat_cty$policyRate_sc        <- safe_scale(dat_cty$policyRate)
  dat_cty$realGDP_sc           <- safe_scale(dat_cty$realGDP)
  dat_cty$usdExchangeRate_sc   <- safe_scale(dat_cty$usdExchangeRate)

  if (!is.finite(sd(dat_cty$headlineInflation, na.rm = TRUE)) || sd(dat_cty$headlineInflation, na.rm = TRUE) == 0) {
    dat_cty$headlineInflation_sc <- seq(-1e-6, 1e-6, length.out = nrow(dat_cty))
  }
  if (!is.finite(sd(dat_cty$policyRate, na.rm = TRUE)) || sd(dat_cty$policyRate, na.rm = TRUE) == 0) {
    dat_cty$policyRate_sc <- seq(-1e-6, 1e-6, length.out = nrow(dat_cty))
  }
  if (!is.finite(sd(dat_cty$realGDP, na.rm = TRUE)) || sd(dat_cty$realGDP, na.rm = TRUE) == 0) {
    dat_cty$realGDP_sc <- seq(-1e-6, 1e-6, length.out = nrow(dat_cty))
  }
  if (!is.finite(sd(dat_cty$usdExchangeRate, na.rm = TRUE)) || sd(dat_cty$usdExchangeRate, na.rm = TRUE) == 0) {
    dat_cty$usdExchangeRate_sc <- seq(-1e-6, 1e-6, length.out = nrow(dat_cty))
  }

  ps_use <- if (cty == "Italy") 4 else 5

  fit <- try(
    vgam(
      y_cty ~
        sm.ps(headlineInflation_sc, ps.int = ps_use) +
        sm.ps(policyRate_sc, ps.int = ps_use) +
        sm.ps(realGDP_sc, ps.int = ps_use) +
        sm.ps(usdExchangeRate_sc, ps.int = ps_use),
      family = dirichlet(),
      data = dat_cty,
      control = vgam.control(Maxit.outer = 200, trace = FALSE)
    ),
    silent = TRUE
  )

  g7_models[[cty]] <- fit

  if (inherits(fit, "try-error")) {
    cat("FAILED:", cty, "\n")
  } else {
    cat("OK:", cty, "| ps.int =", ps_use, "\n")
  }
}


# 4) Japan: special final version
# keep ps.int = 5, stronger adjustment, Maxit.outer = 400

cty <- "Japan"

cat("\n====================\n")
cat("Fitting:", cty, "(special model)\n")

idx <- which(df2_10y$country == cty)
ord <- order(df2_10y$date_m[idx])

dat_jp <- df2_10y[idx[ord], ]
y_jp   <- as.matrix(df2_10y[idx[ord], c("bank", "govern", "growth", "market", "other")])

K_jp <- ncol(y_jp)

y_jp[y_jp < 0] <- 0
y_jp <- y_jp / rowSums(y_jp)

eps_jp <- 2e-3
y_jp <- (y_jp + eps_jp) / (1 + K_jp * eps_jp)

delta_jp <- 0.003
y_jp <- (1 - delta_jp) * y_jp + delta_jp * (1 / K_jp)

y_jp <- y_jp / rowSums(y_jp)
storage.mode(y_jp) <- "double"

dat_jp$headlineInflation_sc <- safe_scale(dat_jp$headlineInflation)
dat_jp$policyRate_sc        <- safe_scale(dat_jp$policyRate)
dat_jp$realGDP_sc           <- safe_scale(dat_jp$realGDP)
dat_jp$usdExchangeRate_sc   <- safe_scale(dat_jp$usdExchangeRate)

if (!is.finite(sd(dat_jp$headlineInflation, na.rm = TRUE)) || sd(dat_jp$headlineInflation, na.rm = TRUE) == 0) {
  dat_jp$headlineInflation_sc <- seq(-1e-6, 1e-6, length.out = nrow(dat_jp))
}
if (!is.finite(sd(dat_jp$policyRate, na.rm = TRUE)) || sd(dat_jp$policyRate, na.rm = TRUE) == 0) {
  dat_jp$policyRate_sc <- seq(-1e-6, 1e-6, length.out = nrow(dat_jp))
}
if (!is.finite(sd(dat_jp$realGDP, na.rm = TRUE)) || sd(dat_jp$realGDP, na.rm = TRUE) == 0) {
  dat_jp$realGDP_sc <- seq(-1e-6, 1e-6, length.out = nrow(dat_jp))
}
if (!is.finite(sd(dat_jp$usdExchangeRate, na.rm = TRUE)) || sd(dat_jp$usdExchangeRate, na.rm = TRUE) == 0) {
  dat_jp$usdExchangeRate_sc <- seq(-1e-6, 1e-6, length.out = nrow(dat_jp))
}

fit_jp <- try(
  vgam(
    y_jp ~
      sm.ps(headlineInflation_sc, ps.int = 5) +
      sm.ps(policyRate_sc, ps.int = 5) +
      sm.ps(realGDP_sc, ps.int = 5) +
      sm.ps(usdExchangeRate_sc, ps.int = 5),
    family = dirichlet(),
    data = dat_jp,
    control = vgam.control(Maxit.outer = 400, trace = FALSE)
  ),
  silent = TRUE
)

g7_models[["Japan"]] <- fit_jp

if (inherits(fit_jp, "try-error")) {
  cat("FAILED: Japan\n")
} else {
  cat("OK: Japan | ps.int = 5 | Maxit.outer = 400\n")
}


# 5) United States: separate model without usdExchangeRate
cty <- "United States"

cat("\n====================\n")
cat("Fitting:", cty, "(separate model)\n")

idx <- which(df2_10y$country == cty)
ord <- order(df2_10y$date_m[idx])

dat_us <- df2_10y[idx[ord], ]
y_us   <- y_adj[idx[ord], , drop = FALSE]

dat_us$headlineInflation_sc <- safe_scale(dat_us$headlineInflation)
dat_us$policyRate_sc        <- safe_scale(dat_us$policyRate)
dat_us$realGDP_sc           <- safe_scale(dat_us$realGDP)

s_hi  <- sd(dat_us$headlineInflation, na.rm = TRUE)
s_pr  <- sd(dat_us$policyRate, na.rm = TRUE)
s_gdp <- sd(dat_us$realGDP, na.rm = TRUE)

if (!is.finite(s_hi) || is.na(s_hi) || s_hi == 0) {
  dat_us$headlineInflation_sc <- seq(-1e-6, 1e-6, length.out = nrow(dat_us))
}
if (!is.finite(s_pr) || is.na(s_pr) || s_pr == 0) {
  dat_us$policyRate_sc <- seq(-1e-6, 1e-6, length.out = nrow(dat_us))
}
if (!is.finite(s_gdp) || is.na(s_gdp) || s_gdp == 0) {
  dat_us$realGDP_sc <- seq(-1e-6, 1e-6, length.out = nrow(dat_us))
}

fit_us <- try(
  vgam(
    y_us ~
      sm.ps(headlineInflation_sc, ps.int = 6) +
      sm.ps(policyRate_sc, ps.int = 6) +
      sm.ps(realGDP_sc, ps.int = 6),
    family = dirichlet(),
    data = dat_us,
    control = vgam.control(Maxit.outer = 200, trace = FALSE)
  ),
  silent = TRUE
)

g7_models[["United States"]] <- fit_us

if (inherits(fit_us, "try-error")) {
  cat("FAILED: United States\n")
} else {
  cat("OK: United States | ps.int = 6 | usdExchangeRate removed\n")
}


# 6) AIC / BIC table
ok_models <- sapply(g7_models, function(x) inherits(x, "vglm"))
ok_countries <- names(g7_models)[ok_models]

ps_used <- ifelse(
  ok_countries == "Italy", 4,
  ifelse(ok_countries == "Japan", 5,
         ifelse(ok_countries == "United States", 6, 5))
)

n_smooth_terms <- ifelse(ok_countries == "United States", 3, 4)

g7_model_compare <- data.frame(
  country = ok_countries,
  n_smooth_terms = n_smooth_terms,
  ps_int_used = ps_used,
  AIC = sapply(g7_models[ok_countries], AIC),
  BIC = sapply(g7_models[ok_countries], BIC)
)

print(g7_model_compare)

```
# In the country-specific models, most countries retained the same four smooth macroeconomic covariates as in the  final model with ps.int = 5, since this provided the best practical balance between visible nonlinear wiggliness, numerical stability, and computational feasibility. Italy was estimated with ps.int = 4 because the four-smooth model became unstable at ps.int = 5, while Japan kept ps.int = 5 after a slightly stronger response adjustment and a larger outer-iteration limit, and the United States was modeled separately with three smooth terms because usdExchangeRate showed essentially no within-country variation and was therefore not identifiable. The output should therefore be interpreted as showing that country-specific models can better capture some medium-run national dynamics, but they are generally harder to fit than the pooled final model, and their AIC/BIC values are most meaningful for model choice within each country rather than for direct cross-country ranking.
```{r G7 model plots }
g7_fitted_data <- list()

for (cty in ok_countries) {
  idx <- which(df2_10y$country == cty)
  ord <- order(df2_10y$date_m[idx])

  dat_cty <- df2_10y[idx[ord], ]
  fit_mu <- fitted(g7_models[[cty]])

  dat_cty$bank_fit   <- fit_mu[, "bank"]
  dat_cty$govern_fit <- fit_mu[, "govern"]
  dat_cty$growth_fit <- fit_mu[, "growth"]
  dat_cty$market_fit <- fit_mu[, "market"]
  dat_cty$other_fit  <- fit_mu[, "other"]

  g7_fitted_data[[cty]] <- dat_cty
}


# Plot function

plot_bank_fit <- function(cty_name) {
  tmp <- g7_fitted_data[[cty_name]]

  plot(
    tmp$date_m, tmp$growth,
    type = "l",
    xlab = "Date",
    ylab = "bank proportion",
    main = paste("Observed vs Fitted (bank) -", cty_name)
  )

  lines(tmp$date_m, tmp$growth_fit, lty = 2)

  legend(
    "topright",
    legend = c("Observed", "Fitted"),
    lty = c(1, 2),
    bty = "n"
  )
}


#  Draw plots

old_par <- par(no.readonly = TRUE)
par(mfrow = c(2, 2), mar = c(4, 4, 3, 1), cex.main = 1.0, cex.lab = 1, cex.axis = 0.9)

for (cty in c("Canada", "France", "Germany", "Italy")) {
  plot_bank_fit(cty)
}

par(old_par)

old_par <- par(no.readonly = TRUE)
par(mfrow = c(2, 2), mar = c(4, 4, 3, 1), cex.main = 1.0, cex.lab = 1, cex.axis = 0.9)

for (cty in c("United Kingdom", "Japan", "United States")) {
  plot_bank_fit(cty)
}

par(old_par)
```
#Although ps.int = 6 is theoretically more flexible, it was computationally infeasible for the single-country models; for example, the Canada model did not finish within about one hour under this specification. In contrast, ps.int = 5 remained practically estimable and was sufficiently flexible to capture the main nonlinear patterns. Additional comparisons for selected countries also indicated that ps.int = 5 performed better than ps.int = 4, so ps.int = 5 was adopted as a practical compromise between model flexibility and computational feasibility.

```{r G7 multi vs single  }

library(dplyr)

# 0. Basic settings

comp <- c("bank", "govern", "growth", "market", "other")

g7_countries <- c(
  "Canada", "France", "Germany", "Italy",
  "Japan", "United Kingdom", "United States"
)

multi_model <- vg_pr5


# 1. Multi-country fitted values

multi_fit_mu <- as.matrix(fitted(multi_model))

multi_fitted_data <- df2_10y %>%
  arrange(country, date_m) %>%
  mutate(
    bank_multi   = multi_fit_mu[, "bank"],
    govern_multi = multi_fit_mu[, "govern"],
    growth_multi = multi_fit_mu[, "growth"],
    market_multi = multi_fit_mu[, "market"],
    other_multi  = multi_fit_mu[, "other"]
  )


# 2. Single-country fitted values

single_fitted_list <- list()

for (cty in g7_countries) {
  idx <- which(df2_10y$country == cty)
  ord <- order(df2_10y$date_m[idx])

  dat_cty <- df2_10y[idx[ord], ]

  fit_mu_single <- as.matrix(fitted(g7_models[[cty]]))

  dat_cty$bank_single   <- fit_mu_single[, "bank"]
  dat_cty$govern_single <- fit_mu_single[, "govern"]
  dat_cty$growth_single <- fit_mu_single[, "growth"]
  dat_cty$market_single <- fit_mu_single[, "market"]
  dat_cty$other_single  <- fit_mu_single[, "other"]

  single_fitted_list[[cty]] <- dat_cty
}

single_fitted_data <- bind_rows(single_fitted_list)


# 3. Merge observed, multi-country, and single-country values

compare_df <- multi_fitted_data %>%
  select(
    country, date_m,
    bank, govern, growth, market, other,
    bank_multi, govern_multi, growth_multi, market_multi, other_multi
  ) %>%
  left_join(
    single_fitted_data %>%
      select(
        country, date_m,
        bank_single, govern_single, growth_single, market_single, other_single
      ),
    by = c("country", "date_m")
  ) %>%
  arrange(country, date_m)


# 4. Plot function: one country = 5 anchor plots

plot_country_5anchors <- function(cty_name, data_plot = compare_df) {

  tmp <- data_plot %>%
    filter(country == cty_name) %>%
    arrange(date_m)

  old_par <- par(no.readonly = TRUE)
  on.exit(par(old_par))

  par(
    mfrow = c(3, 2),
    mar = c(4.2, 4.2, 2.8, 1.2),
    oma = c(0.5, 0.5, 2.2, 0.5),
    cex.main = 1.05,
    cex.lab = 0.95,
    cex.axis = 0.9
  )

  # bank
  yr <- range(c(tmp$bank, tmp$bank_multi, tmp$bank_single), na.rm = TRUE)
  plot(tmp$date_m, tmp$bank,
       type = "l", lwd = 1, lty = 3,
       xlab = "Date", ylab = "Proportion",
       main = "bank", ylim = yr)
  lines(tmp$date_m, tmp$bank_multi,  lwd = 2.4, lty = 1)
  lines(tmp$date_m, tmp$bank_single, lwd = 2.4, lty = 2)
  legend("topright",
         legend = c("Observed", "Multi-country", "Single-country"),
         lty = c(3, 1, 2), lwd = c(1, 2.4, 2.4),
         bty = "n", cex = 0.82)

  # govern
  yr <- range(c(tmp$govern, tmp$govern_multi, tmp$govern_single), na.rm = TRUE)
  plot(tmp$date_m, tmp$govern,
       type = "l", lwd = 1, lty = 3,
       xlab = "Date", ylab = "Proportion",
       main = "govern", ylim = yr)
  lines(tmp$date_m, tmp$govern_multi,  lwd = 2.4, lty = 1)
  lines(tmp$date_m, tmp$govern_single, lwd = 2.4, lty = 2)
  legend("topright",
         legend = c("Observed", "Multi-country", "Single-country"),
         lty = c(3, 1, 2), lwd = c(1, 2.4, 2.4),
         bty = "n", cex = 0.82)

  # growth
  yr <- range(c(tmp$growth, tmp$growth_multi, tmp$growth_single), na.rm = TRUE)
  plot(tmp$date_m, tmp$growth,
       type = "l", lwd = 1, lty = 3,
       xlab = "Date", ylab = "Proportion",
       main = "growth", ylim = yr)
  lines(tmp$date_m, tmp$growth_multi,  lwd = 2.4, lty = 1)
  lines(tmp$date_m, tmp$growth_single, lwd = 2.4, lty = 2)
  legend("topright",
         legend = c("Observed", "Multi-country", "Single-country"),
         lty = c(3, 1, 2), lwd = c(1, 2.4, 2.4),
         bty = "n", cex = 0.82)

  # market
  yr <- range(c(tmp$market, tmp$market_multi, tmp$market_single), na.rm = TRUE)
  plot(tmp$date_m, tmp$market,
       type = "l", lwd = 1, lty = 3,
       xlab = "Date", ylab = "Proportion",
       main = "market", ylim = yr)
  lines(tmp$date_m, tmp$market_multi,  lwd = 2.4, lty = 1)
  lines(tmp$date_m, tmp$market_single, lwd = 2.4, lty = 2)
  legend("topright",
         legend = c("Observed", "Multi-country", "Single-country"),
         lty = c(3, 1, 2), lwd = c(1, 2.4, 2.4),
         bty = "n", cex = 0.82)

  # other
  yr <- range(c(tmp$other, tmp$other_multi, tmp$other_single), na.rm = TRUE)
  plot(tmp$date_m, tmp$other,
       type = "l", lwd = 1, lty = 3,
       xlab = "Date", ylab = "Proportion",
       main = "other", ylim = yr)
  lines(tmp$date_m, tmp$other_multi,  lwd = 2.4, lty = 1)
  lines(tmp$date_m, tmp$other_single, lwd = 2.4, lty = 2)
  legend("topright",
         legend = c("Observed", "Multi-country", "Single-country"),
         lty = c(3, 1, 2), lwd = c(1, 2.4, 2.4),
         bty = "n", cex = 0.82)

  plot.new()

  mtext(cty_name, outer = TRUE, cex = 1.5, font = 2)
}




# 5. Export all G7 countries to one PDF

pdf("G7_5anchor_observed_multi_single_clean.pdf", width = 12, height = 8.8)

for (cty in g7_countries) {
  plot_country_5anchors(cty)
}

dev.off()

```
# Increasing Maxit.outer from 200 to 400 leads to some improvement, but the change is not dramatic. For France, the fitted curve becomes slightly more stable and responsive to medium-run movements, while for Germany the improvement is more visible, especially around broader trend changes. This suggests that a higher outer-iteration limit can help some country-specific models reach a better solution, but the benefit is uneven across countries.

```{r G7 multi vs single coefficient }


#  Basic settings

comp <- c("bank", "govern", "growth", "market", "other")

g7_countries <- c(
  "Canada", "France", "Germany", "Italy",
  "Japan", "United Kingdom", "United States"
)

multi_model <- vg_pr5


# 1. Multi-country fitted values

multi_fit_mu <- as.matrix(fitted(multi_model))

multi_fitted_data <- df2_10y %>%
  arrange(country, date_m) %>%
  mutate(
    bank_multi   = multi_fit_mu[, "bank"],
    govern_multi = multi_fit_mu[, "govern"],
    growth_multi = multi_fit_mu[, "growth"],
    market_multi = multi_fit_mu[, "market"],
    other_multi  = multi_fit_mu[, "other"]
  )


# 2. Single-country fitted values

single_fitted_list <- list()

for (cty in g7_countries) {
  idx <- which(df2_10y$country == cty)
  ord <- order(df2_10y$date_m[idx])

  dat_cty <- df2_10y[idx[ord], ]

  fit_mu_single <- as.matrix(fitted(g7_models[[cty]]))

  dat_cty$bank_single   <- fit_mu_single[, "bank"]
  dat_cty$govern_single <- fit_mu_single[, "govern"]
  dat_cty$growth_single <- fit_mu_single[, "growth"]
  dat_cty$market_single <- fit_mu_single[, "market"]
  dat_cty$other_single  <- fit_mu_single[, "other"]

  single_fitted_list[[cty]] <- dat_cty
}

single_fitted_data <- bind_rows(single_fitted_list)


# 3. Merge observed, multi-country, and single-country values

compare_df <- multi_fitted_data %>%
  select(
    country, date_m,
    bank, govern, growth, market, other,
    bank_multi, govern_multi, growth_multi, market_multi, other_multi
  ) %>%
  left_join(
    single_fitted_data %>%
      select(
        country, date_m,
        bank_single, govern_single, growth_single, market_single, other_single
      ),
    by = c("country", "date_m")
  ) %>%
  arrange(country, date_m)


# 4. Error functions

mae_fun <- function(obs, fit) {
  mean(abs(obs - fit), na.rm = TRUE)
}

rmse_fun <- function(obs, fit) {
  sqrt(mean((obs - fit)^2, na.rm = TRUE))
}


# 5. Composition-level comparison:

error_comp_list <- list()

for (cty in g7_countries) {

  tmp <- compare_df %>%
    filter(country == cty) %>%
    arrange(date_m)

  # Observed composition values
  obs_mat <- as.matrix(tmp[, comp])

  # Multi-country fitted composition values
  fit_multi_mat <- as.matrix(tmp[, paste0(comp, "_multi")])

  # Single-country fitted composition values
  fit_single_mat <- as.matrix(tmp[, paste0(comp, "_single")])

  # Flatten matrices into vectors
  obs_vec <- as.vector(obs_mat)
  fit_multi_vec <- as.vector(fit_multi_mat)
  fit_single_vec <- as.vector(fit_single_mat)

  error_comp_list[[cty]] <- data.frame(
    country = cty,
    multi_MAE = mae_fun(obs_vec, fit_multi_vec),
    multi_RMSE = rmse_fun(obs_vec, fit_multi_vec),
    single_MAE = mae_fun(obs_vec, fit_single_vec),
    single_RMSE = rmse_fun(obs_vec, fit_single_vec)
  )
}

error_composition_country <- bind_rows(error_comp_list) %>%
  arrange(country)


# 6. Print results

print(error_composition_country)
```
